The LSST DESC 3x2pt Tomography Optimization Challenge

### 2021-01-29 paper/paper.tex
Add PCACluster

\subsection{ {\sc PCACluster} }

>>>>

\subsection{ {\sc PCACluster} }
This entry was submitted by author DG.

Principal Component Analysis (PCA) reduces the dimensionality of a dataset by
calculating eigenvectors that align with the principal axes of variation in the
data \citep{doi:10.1098}. {\sc PCACluster} uses PCA to reduce the flux \& color
data set to three dimensions.

To generate the PCA dimensions used by this method the PCA algorithm takes in
the $r$-band flux and $ri$ and $iz$ colors and outputs three principal
components.  If the data set provides the $g$-band, the algorithm also uses the
$gr$ color when determining eigenvectors. Tomographic bins are then determined
in this three dimensional space using a unique clustering algorithm where each
observation is assigned to a bin by determining which cluster centroid is
closest to that observation.

In order to determine optimal centroid positions, clustering is framed as a
gradient descent problem that seeks to maximize the requested metric: either the
SNR or the FoM. This approach of identifying clusters using gradient descent is
inspired by the similar approach to K-means clustering in
\cite{NIPS1994_a1140a3d}.

This entry uses {\sc jax\_cosmo} to take the derivative of the requested metric
and classification function combination with respect to the centroid locations
to calculate gradients and weight updates \citep{jax-cosmo}.  The speed of
training PCACluster is directly proportional to the number of requested
centroids (the number of requested tomographic bins) and the amount of training
data used.
